Yield bias

### 2015-10-01 Rmd/yield-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

```{r, eval=FALSE}
library(SAcropland)
library(xtable)
library(RColorBrewer)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")

@@

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100)

@@

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(gti, stack(lc_list)), sumna)
namask[namask > 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
```

@@

@@

# tst <- gti_agg$f5$gti - lc_agg$f5$sa30
# plot(tst * 100)

#pct_diff <- function(x, y) (x - y) / x * 100

cf_pct_diff <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- gti_agg[[y]][[1]] - lc_agg[[y]][[x]]
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff) <- snms

# disaggregate for plotting
namaskc <- crop(namask, gtic)
lev <- names(cf_pct_diff[[1]])
disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, crop(namask, gtic))

@@

# redo NA mask because of some slight shifts
namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
namask[namask > 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop
# sumna <- function(x, na.rm = na.rm) sum(x, na.rm = na.rm)
# amsk <- aggregate_rast_list(fact, list(namask2), fun = sumna)
# amsk <- lapply(amsk, function(x) x[[1]])  # unlist on inner loop

@@

# not implausible that someone would do a direct aggregation of these yield 
# values, without weighting by crop distribution, so this incorrect variant 
# would be: 
cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg

# third variant - aggregate and remove zeros
# first define a function that removes zero yield areas when calculating mean 
# yield--this will allow for aggregation of yields only in those areas having 
# maize, according to Monfreda et al methods
nzmean <- function(x, na.rm) {
  x <- x[(x > 0) & !is.na(x)]
  if(length(x) == 0) {
    m <- 0
  } else if(length(x) > 0) {
    m <- mean(x, na.rm = na.rm)
  }
  return(m)
}
cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg

@@

yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
## check values
# check yields
for(i in lev[-1]) {
  print(paste("factor", i))
  for(j in asnms) {
    print(paste("...", j))
    for(k in 1:3) {
      #ck <- cellStats(cyld_agg$f1$gti, nzmean)
      print(paste("...... type", k, "agg"))
      ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
      ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
      w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
      x <- values(yat[[k]][[i]][[j]])  # yield values
      ind <- which(x > 0)  # which are non-zero
      x <- x[ind]  # select non zeros
      w <- w[ind]  # select non-zeros
      #w <- w / max(w)
      v <- weighted.mean(x, w)
      d <- round((ck - v) / ck, 2)
      d2 <- round((ck2 - v) / ck2, 2)
      print(c(d, d2))
    }
  }  # type 1 and type 3 aggregations consistent between datasets, across scales
}
rm(yat)

@@

# with 0-removed yield aggregation (type 3 yield aggregation)
cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
  p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
    (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
  })
  named_out(p, asnms)
})
names(cpagg3) <- names(agg_res)

@@

# check production estimates for consistency
pat <- list(cpagg1, cpagg3, cpaggII)
ptype <- c("1", "3", "II")
# check production estimates
for(i in lev[-1]) {
  print(paste("factor", i))
  for(j in asnms) {
    print(paste("...", j))
    for(k in 1:3) {
      print(paste("...... type", ptype[k], "agg"))
      ck <- cellStats(cprod[[j]], sum)  # weighted by fraction
      ck2 <- cellStats(cprod$gti, sum) # weighted by fraction
      v <- cellStats(pat[[k]][[i]][[j]], sum)
      d <- round((ck - v) / ck, 2)
      d2 <- round((ck2 - v) / ck2, 2)
      print(c(d, d2))
    }
  }  # all types consistent across scales
}

@@

# Mean production in producing cells at different levels of aggregation
mup <- sapply(lev, function(i) {
  msk <- cpaggII[[i]]$gti > 0
  msk[msk == 0] <- NA
  cpmsk <- msk * cpaggII[[i]]$gti
  cellStats(cpmsk, mean)
})
# sapply(lev, function(i) {
#   msk <- cpagg1[[i]]$gti > 0
#   msk[msk == 0] <- NA
#   cpmsk <- msk * cpagg1[[i]]$gti
#   cellStats(cpmsk, mean)
# })

# mean yield in producing cells at different levels of aggregation
muy <- sapply(lev, function(i) {
  w <- values(carea_agg[[i]]$gti)  # aggregated crop fraction
  x <- values(cyld_agg[[i]]$gti)  # yield values
  ind <- which(x > 0)  # which are non-zero
  x <- x[ind]  # select non zeros
  w <- w[ind]  # select non-zeros
  v <- weighted.mean(x, w, na.rm = TRUE, )
  v
})

@@

# Production differences
ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
list1 <- lapply(cpaggII, function(x) x[[1]])
list2 <- lapply(cpagg1, function(x) x[-1])
p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg
list2 <- lapply(cpagg3, function(x) x[-1])
p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
list2 <- lapply(cpaggII, function(x) x[-1])
p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
# pmu <- as.data.table.raster(cpagg3$f1$globmu)[layer > 0, mean(layer)]

@@

# type 2 (non-zero)
list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2

# convert them to percent differences relative to mean pixel-wise production
# here only for map plotting.
# plot(y_diff2$sa30$f25$gti)
pdiffs <- lapply(list(p_diff1, p_diff2, p_diff3), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / mup[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
names(pdiffs) <- paste0("pd", 1:3)

ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / muy[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
names(ydiffs) <- paste0("yd", 1:2)

# disaggregate for plotting
lev <- names(y_diff2[[1]])
p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)
y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
stats_pd1 <- dstats(p_diff1d)
stats_pd2 <- dstats(p_diff2d)
stats_pd3 <- dstats(p_diff3d)
stats_yd1 <- dstats(y_diff1d)
stats_yd2 <- dstats(y_diff2d)
# cbind(stats_yd1$geow[, 4], stats_yd2$geow[, 4])
```
<a href="#top">Back to top</a>

@@

fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
          paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
# legtext <- c(rep("tons", 3), rep("tons/ha", 2))
legtext <- rep("% Difference", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(p_diff1d[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")

@@

From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates. There are three types of yield aggregation that can be done:

@@

@@

# first check aggregation to make yield aggregation variants have common area,
# so that just one lc_union is needed
for(i in lev) {
  for(j in snms) {
    print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
    #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
  }
}  # all zeros

snms <- names(p_diff1)
lcu <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(carea_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(carea_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev
# plot(lc_union$f50$modmu)
# plot(gti_agg$f50$gti)

@@

# Spatial bias stats calculation on yield and production differences, not on
# percentages
a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, p_diff1, wma, "mua", "bias", TRUE)
d <- bias_stats_list(bins, awgts, lcu, p_diff1, sum, "sum", "bias", FALSE)
pd1_st <- rbind(a, b, d)
a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
pd2_st <- rbind(a, b, d)
a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
pd3_st <- rbind(a, b, d)
a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, y_diff1, wma, "mua", "bias", TRUE)
yd1_st <- rbind(a, b)
a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)      
yd2_st <- rbind(a, b)

# Output statistics
p1mu <- extract_stat(lev, snms, "all", "mu", "bias", pd1_st)  # identical to p3
p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)
y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)

# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
i <- y_diff1 #pdiff_3 # p_diff1
jdt <- yd1_st # pd3_st # pd1_st
bv <- "mu"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- i[[x]][[y]][[1]]  
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(jdt[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

>>>>

```{r, eval=FALSE}
library(SAcropland)
library(xtable)
library(RColorBrewer)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
# load(full_path(p_data, "yield-bias.rda"))

@@

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
# cellStats(!is.na(lc_list$sa30), sum)

@@

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
# namask <- calc(stack(stack(gti), stack(lclist)), sumna)
# namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
```

@@

# set up masks and weights for assessing output bias and absolute errors
namask <- calc(stack(gtic, lc_agg$f1), sumna)
# cellStats(!is.na(gtic), sum); cellStats(!is.na(lc_agg$f1$sa30), sum)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(namask), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop

@@

# tst <- gti_agg$f5$gti - lc_agg$f5$sa30
# plot(tst * 100)
#pct_diff <- function(x, y) (x - y) / x * 100

# calculate differences between adjusted fraction and reference
cf_pct_diff <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- gti_agg[[y]][[1]] - lc_agg[[y]][[x]]
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff) <- snms

# same, but as percent
cf_pct_diff2 <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- (gti_agg[[y]][[1]] - lc_agg[[y]][[x]]) * 100
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff2) <- snms

# cropland cover bins, for looking at error as a function of cover, based on 
# original extent of gti data, for ease
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg, function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})

# snms <- names(p_diff1)
# lcnu <- lapply(lev, function(x) {
#   lcb <- lapply(snms, function(y) {
#     all_gt0 <- !is.na(bins[[x]])
#     all_gt0
#   })
#   named_out(lcb, snms)
# })
# names(lcnu) <- lev
lev <- names(cf_pct_diff[[1]])
snms <- names(p_diff1)
lcu <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

# calculate means
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
a <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wma, "mua", "bias", TRUE)
# b <- bias_stats_list(bins, awgts, lcnu, cf_pct_diff2, wma, "mua", "bias", TRUE)
# b[ol == "f1" & il == "modmu" & bin == "all"]

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
bv <- "mua"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(cf_pct_diff2[[x]][[y]])
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

lcfmua <- extract_stat(lev, snms, "all", "mua", "bias", b)
lcfmua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", b) 
# b[ol == "f10" & il == "globmu" & bin == "all"]

svec <- c("mua.nm", "mua")
aa <- c("Country", "Agricultural")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

lcf_out <- do.call(rbind, lapply(1:length(svec), function(i) {
  a <- do.call(rbind, lapply(snms, function(x) {
    aa <- sapply(lev, function(y) {
      b[ol == y & il == x & bin == "all" & bvals == svec[i], bias]
    })
    named_out(aa, paste(c(1, fact), "km"))
  }))
  cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
}))

caption <- paste("Mean absolute errors (MAE) in statistically constrained", 
                 "cropland maps, calculated over the entire country, and for", 
                 "the union of agricultural regions (cropland $>$ 0) defined ", 
                 "by the reference and corresponding adjusted cropland",
                 "maps")    
lcfout_xtab <- xtable(lcf_out, digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", file = "paper/figures/cropadj-mae.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

# disaggregate for plotting
# namaskc <- crop(namask, gtic)
disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, namask)

@@

# redo NA mask because of some slight shifts
namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop
# sumna <- function(x, na.rm = na.rm) sum(x, na.rm = na.rm)
# amsk <- aggregate_rast_list(fact, list(namask2), fun = sumna)
# amsk <- lapply(amsk, function(x) x[[1]])  # unlist on inner loop
```
<a href="#top">Back to top</a>

@@

# not implausible that someone would do a direct aggregation of these yield 
# values, without weighting by crop distribution, so this incorrect variant 
# would be: 
# cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg

# third variant - aggregate and remove zeros
# first define a function that removes zero yield areas when calculating mean 
# yield--this will allow for aggregation of yields only in those areas having 
# maize, according to Monfreda et al methods
# nzmean <- function(x, na.rm) {
#   x <- x[(x > 0) & !is.na(x)]
#   if(length(x) == 0) {
#     m <- 0
#   } else if(length(x) > 0) {
#     m <- mean(x, na.rm = na.rm)
#   }
#   return(m)
# }
# cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg

@@

# yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
# ## check values
# # check yields
# for(i in lev[-1]) {
#   print(paste("factor", i))
#   for(j in asnms) {
#     print(paste("...", j))
#     for(k in 1:3) {
#       #ck <- cellStats(cyld_agg$f1$gti, nzmean)
#       print(paste("...... type", k, "agg"))
#       ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
#       ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
#       w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
#       x <- values(yat[[k]][[i]][[j]])  # yield values
#       ind <- which(x > 0)  # which are non-zero
#       x <- x[ind]  # select non zeros
#       w <- w[ind]  # select non-zeros
#       #w <- w / max(w)
#       v <- weighted.mean(x, w)
#       d <- round((ck - v) / ck, 2)
#       d2 <- round((ck2 - v) / ck2, 2)
#       print(c(d, d2))
#     }
#   }  # type 1 and type 3 aggregations consistent between datasets, across scales
# }
# rm(yat)

@@

# with 0-removed yield aggregation (type 3 yield aggregation)
# cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
#   p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
#     (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
#   })
#   named_out(p, asnms)
# })
# names(cpagg3) <- names(agg_res)

@@

# check production estimates for consistency
# pat <- list(cpagg1, cpagg3, cpaggII)
# ptype <- c("1", "3", "II")
pat <- list(cpagg1, cpaggII)
ptype <- c("1", "II")

# check production estimates
for(i in lev[-1]) {
  print(paste("factor", i))
  for(j in asnms) {
    print(paste("...", j))
    for(k in 1:length(ptype)) {
      print(paste("...... type", ptype[k], "agg"))
      ck <- cellStats(cprod[[j]], sum)  # weighted by fraction
      ck2 <- cellStats(cprod$gti, sum) # weighted by fraction
      v <- cellStats(pat[[k]][[i]][[j]], sum)
      d <- round((ck - v) / ck, 2)
      d2 <- round((ck2 - v) / ck2, 2)
      print(c(d, d2))
    }
  }  # all types consistent across scales
}

@@

# Mean production in producing cells at different levels of aggregation
mup <- sapply(lev, function(i) {
  msk <- cpagg1[[i]]$gti > 0
  #msk <- cpaggII[[i]]$gti > 0
  msk[msk == 0] <- NA
  cpmsk <- msk * cpagg1[[i]]$gti
  cellStats(cpmsk, mean)
})
# sapply(lev, function(i) {
#   msk <- cpagg1[[i]]$gti > 0
#   msk[msk == 0] <- NA
#   cpmsk <- msk * cpagg1[[i]]$gti
#   cellStats(cpmsk, mean)
# })

# mean yield in producing cells at different levels of aggregation
muy <- sapply(lev, function(i) {
  w <- values(carea_agg[[i]]$gti)  # aggregated crop fraction
  x <- values(cyld_agg[[i]]$gti)  # yield values
  ind <- which(x > 0)  # which are non-zero
  x <- x[ind]  # select non zeros
  w <- w[ind]  # select non-zeros
  v <- weighted.mean(x, w, na.rm = TRUE, )
  v
})  # yields across scales and datasets all 3.4-3.5 kg/ha

@@

# Production differences
ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
list1 <- lapply(cpagg1, function(x) x[1])
list2 <- lapply(cpagg1, function(x) x[-1])
p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg
# list2 <- lapply(cpagg3, function(x) x[-1])
# p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
# list2 <- lapply(cpaggII, function(x) x[-1])
# p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
# pmu <- as.data.table.raster(cpagg3$f1$globmu)[layer > 0, mean(layer)]

@@

# type 2 (non-zero)
# list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
# y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2

# convert them to percent differences relative to mean pixel-wise production
# here only for map plotting.
pdiffs <- lapply(list(p_diff1), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / mup[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
# names(pdiffs) <- paste0("pd", 1:3)
names(pdiffs) <- "pd1"

# ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
ydiffs <- lapply(list(y_diff1), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / muy[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
# names(ydiffs) <- paste0("yd", 1:2)
names(ydiffs) <- "yd1"

# disaggregate for plotting
lev <- names(y_diff1[[1]])
p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
# p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
# p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)
# y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
stats_pd1 <- dstats(p_diff1d)
# stats_pd2 <- dstats(p_diff2d)
# stats_pd3 <- dstats(p_diff3d)
stats_yd1 <- dstats(y_diff1d)
# stats_yd2 <- dstats(y_diff2d)
# cbind(stats_yd1$geow[, 4], stats_yd2$geow[, 4])
```
<a href="#top">Back to top</a>

@@

# fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
#           paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
fnms <- c("prod_bias_map_1.pdf", "yld_bias_map_1.pdf")  # output names
# legtext <- c(rep("tons", 3), rep("tons/ha", 2))
legtext <- rep("% Difference", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(p_diff1d[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")

@@

From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates. There are three types of yield aggregation that can be done (most commented out now after initial checking):

@@

Only Types 1 and A are included in further investigation. The other variants are left as accuracy checks.

@@

# first check aggregation to make yield aggregation variants have common area,
# so that just one lc_union is needed
# for(i in lev) {
#   for(j in snms) {
#     print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
#     #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
#   }
# }  # all zeros

snms <- names(p_diff1)
lcu <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(carea_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(carea_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev
# plot(lcu$f50$modmu)
# plot(gti_agg$f50$gti)

@@

# Spatial bias stats calculation on yield and production differences, not on
# percentages
a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, p_diff1, wma, "mua", "bias", TRUE)
d <- bias_stats_list(bins, awgts, lcu, p_diff1, sum, "sum", "bias", FALSE)
pd1_st <- rbind(a, b, d)
# a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
# b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
# d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
# pd2_st <- rbind(a, b, d)
# a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
# b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
# d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
# pd3_st <- rbind(a, b, d)
a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, y_diff1, wma, "mua", "bias", TRUE)
yd1_st <- rbind(a, b)
# a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
# b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)    
# yd2_st <- rbind(a, b)

# Output statistics
p1mu <- extract_stat(lev, snms, "all", "mu", "bias", pd1_st)  # identical to p3
p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
# p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
# p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
# p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
# p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
# p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
# p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
# p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)
# y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
# y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)

# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
i <- y_diff1 #pdiff_3 # p_diff1
jdt <- yd1_st # pd3_st # pd1_st
bv <- "mua" # bv <- "mu"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(i[[x]][[y]][[1]])
  # l3 <- i[[x]][[y]][[1]]
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(jdt[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check
